---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->

[![Travis build status](https://travis-ci.org/makinin/rdbhapi.svg?branch=master)](https://travis-ci.org/makinin/rdbhapi)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/makinin/rdbhapi?branch=master&svg=true)](https://ci.appveyor.com/project/makinin/rdbhapi)
[![Codecov test coverage](https://codecov.io/gh/makinin/rdbhapi/branch/master/graph/badge.svg)](https://codecov.io/gh/makinin/rdbhapi?branch=master)

<!-- badges: end -->

R interface for [DBH-API](https://dbh.nsd.uib.no/tjenester.action) open data access.

## Installation

You can install the released version of rdbhapi from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("rdbhapi")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("makinin/rdbhapi")

```
Token users have to obtain token in R using function `dbh_api_token(brukernavn, passord)` and
after that they can use R with all token benefits.

## Example

DBH-API contents are in table

```{r content}
library(rdbhapi)
dbh_tabell(1)
```

Get the whole table in R format:

```{r example}
library(rdbhapi)
dbh_tabell(211)
```


Multiple choice query:

```{r query}
dbh_tabell(142, filters = list("Årstall" = c("top","5"),Utvekslingsavtale = "ERASMUS+", 
Type = "NORSK", "Nivåkode" = "*"),exclude = c("Nivåkode" = "FU"), group_by = "Årstall")
```
Meta data for data table  
```{r metadata}
dbh_metadata(142)
```

