---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->


[![Codecov test coverage](https://codecov.io/gh/mni-hkdir/rdbhapi/branch/master/graph/badge.svg)](https://codecov.io/gh/makinin/rdbhapi?branch=master)

[![R-CMD-check](https://github.com/mni-hkdir/rdbhapi/workflows/R-CMD-check/badge.svg)](https://github.com/makinin/rdbhapi/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/rdbhapi)](https://CRAN.R-project.org/package=rdbhapi)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)

R interface for [DBH-API](https://dbh.hkdir.no/dbhapiklient/) open data access.

## Installation



You can install development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("mni-hkdir/rdbhapi")

```
To authenticate with the DBH-API, you need to provide a valid token using your SSO ID and SSO Secret. These credentials can be securely defined as environment variables in one of the following ways:
 Option 1: Using `.Renviron` File

The `.Renviron` file is a hidden file used by R to store environment variables.
You can create or edit this file in your home directory or in your R project folder.
Open the `.Renviron` file in a text editor or within R using:

```{r
file.edit("~/.Renviron")
```
Add  `dbhapi_sso_id` = "your_sso_id" and `dbhapi_sso_secret` = "your_sso_secret"
in `.Renviron` file and save the file. 

After saving, restart your R session to load the updated `.Renviron` file.

Once defined, these variables can be accessed in R using `Sys.setenv`:

```{r
Sys.getenv("dbhapi_sso_id")
Sys.getenv("dbhapi_sso_secret")
```
 Option 2: Using `Sys.setenv`
Alternatively, you can define the token dynamically during your R session using the `Sys.setenv` function:

```{r
Sys.setenv(dbhapi_sso_id = "your_sso_id")
Sys.setenv(dbhapi_sso_secret = "your_sso_secret")

```

## Example

DBH-API contents are in table

```{r content}
library(rdbhapi) 
dbh_data(1)
```

##Get the whole table in R format:

```{r example , warning=FALSE, message=FALSE, results='asis',echo=TRUE, eval=TRUE}
library(knitr)
library(rdbhapi)
institusjoner <- dbh_data(211)
kable(head(institusjoner))
```

##Download the entire table with ID 142:

```{r utvekslingstudenter, warning=FALSE, message=FALSE, results='asis',echo=TRUE, eval=TRUE}
library(knitr)
library(rdbhapi)
utvekslingstudenter <- dbh_data(142)
kable(head(utvekslingstudenter))
```


## Download filtered  data for table with ID 211:

```{r institusjoner_filter, warning=FALSE, message=FALSE, results='asis', echo=TRUE, eval=TRUE}
library(knitr)
library(rdbhapi)
institusjoner_filter <- dbh_data(
  211,
  filters = list(
    "Institusjonskode" = c("top", "5")
  )
)
kable(head(institusjoner_filter))

```

