---
title: "Tutorial / vignette for DBH R package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R_dbh_api}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## R interface for NSD-Database for høgre utdanning(DBH) open data

This R package [rdbhapi](https://github.com/makinin/rdbhapi) provides tools to access [DBH-API](https://dbh.nsd.uib.no/tjenester.action).

## Installation

You can install the released version of rdbhapi from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("rdbhapi")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("makinin/rdbhapi")
```
```{r setup}
library(rdbhapi)
```
Package consists of the following functions

```{r, echo=FALSE,comment=NA}
cat(paste0(library(help = "rdbhapi")$info[[2]], collapse = "\n"))
```

## Downloading data

The package supports two methods for downloading data from DBH-API: downloading of bulk data and sending query to DBH-API.
Both methods are available in package main function `dbh_tabell()`. If function has only tabele id, than the whole table is 
downloaded from the bulk download data. If we use `filters`, `group_by`, `sort_by` than query will be sent in JSON format to DBH-API.

Get the whole table in R format:

```{r rdbhapi, warning=FALSE, message=FALSE, results='asis'}
institusjoner <- dbh_tabell(211)

```

You can get data using a query syntax which is similar to those in DBH-API:

```{r , warning=FALSE, message=FALSE, results='asis',echo=TRUE, eval=TRUE}
library(knitr)
erasmus_plus <- dbh_tabell(142, filters = list("Årstall" = c("top","5"),Utvekslingsavtale = "ERASMUS+", 
Type = "NORSK", "Nivåkode" = "*"),exclude = c("Nivåkode" = "FU"), group_by = "Årstall")
kable(head(erasmus_plus))
```

Token users have to first obtain token in R using function `dbh_api_token()` after that they can use R with all token benefits.

