---
title: "Tools for downloading DBH-API data "
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



This R package [rdbhapi](https://github.com/makinin/rdbhapi) provides tools to access [DBH-API](https://dbh.nsd.uib.no/tjenester.action).


## Installation

You can install the released version of rdbhapi from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("rdbhapi")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("makinin/rdbhapi")
```
```{r setup}
library(rdbhapi)
```
Package consists of the following functions

```{r, echo=FALSE,comment=NA}
cat(paste0(library(help = "rdbhapi")$info[[2]], collapse = "\n"))
```

## Find data

DBH-API contents are in table 

```{r content}
library(rdbhapi)
dbh_data(1)
```


## Downloading data

The package supports two methods for downloading data from DBH-API: 
downloading of bulk data and sending query to DBH-API.
Both methods are available in package main function `dbh_tabell()`. 
If function has only `tabell_id`, than the whole table is 
downloaded from the bulk download data. If we use `filters`, `group_by`, `sort_by` 
than query will be sent in JSON format to DBH-API.

Get the whole table in R format:

```{r rdbhapi, warning=FALSE, message=FALSE, results='asis'}
institusjoner <- dbh_data(211)
```


#### Example for multiple choice query 

Or you can get dataset using query by defining `filters`, `group_by`, `sort_by`, `exclude`  argument. Filter
should be named list, where names corresponds to variable names in DBH-API and
values are vectors corresponding to filter type and filter values. Filter type are the same as in DBH-API: item, all, between, lessthan, greaterthan. For aggregate datasets we need to define `group_by` variables. 


```{r erasmus_plus, warning=FALSE, message=FALSE, results='asis',echo=TRUE, eval=TRUE}
library(knitr)
erasmus_plus <- dbh_data(142, filters = list("Årstall" = c("top","5"),Utvekslingsavtale = "ERASMUS+", 
Type = "NORSK", "Nivåkode" = "*"),exclude = c("Nivåkode" = "FU"), group_by = "Årstall")
kable(head(erasmus_plus))
```



## Token facility 
Token can be defined by placing login credentials in the environment variables `dbhapi_sso_id` and `dbhapi_sso_secret`
in the `.Renviron` file before starting R or by using `Sys.setenv`.

## Session info
This vignette was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```
