---
title: "Interface to 'DBH-API'"
author: "Marija Ninic"
output:
  rmarkdown::html_vignette:
    includes:
      in_header: favicon.html
vignette: >
  %\VignetteIndexEntry{Interface to 'DBH-API'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



[![Codecov test coverage](https://codecov.io/gh/mni-hkdir/rdbhapi/branch/master/graph/badge.svg)](https://codecov.io/gh/mni-hkdir/rdbhapi?branch=master)

[![R-CMD-check](https://github.com/mni-hkdir/rdbhapi/workflows/R-CMD-check/badge.svg)](https://github.com/mni-hkdir/rdbhapi/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/rdbhapi)](https://CRAN.R-project.org/package=rdbhapi)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)





```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```





The [rdbhapi] package provides a user-friendly R interface to interact with 
    the [DBH-API](https://dbh.hkdir.no/dbhapiklient/) for open data access. 
    This API offers comprehensive data on higher education statistics in Norway, 
    enabling seamless access to datasets related to institutions, programs, 
    students, and more. 


## Installation



And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mni-hkdir/rdbhapi")
```
```{r setup}
library(rdbhapi)
```
Package consists of the following functions

```{r, echo=FALSE,comment=NA}
cat(paste0(library(help = "rdbhapi")$info[[2]], collapse = "\n"))
```

## Find data

DBH-API contents are in table 

```{r content}
library(rdbhapi)
dbh_data(1)
```

Retrieve metadata about the DBH-API tables


```{r metadata}
library(rdbhapi)
dbh_data(2)
```

## Downloading data

The rdbhapi package provides two efficient methods for downloading data from the DBH-API:

1. Bulk Data Download: Retrieve an entire table directly.
2. Custom Queries: Send a JSON query to the DBH-API with filters, grouping, or sorting options.
Both methods are implemented in the package's main function, dbh_data(). The method used depends on the arguments passed to the function:

* Bulk Download: If only the tabell_id argument is provided, the entire table is downloaded.
* Querying: If arguments like filters, group_by, or sort_by are included, the function sends a custom JSON query to the DBH-API.  The query is defined by the following arguments in the dbh_data() function:

* filters: A named list where names correspond to variable names in the DBH-API, and values are vectors      specifying filter types and values.Filter types are: item, all, between, lessthan, greaterthan. 
*group_by: Variables to group by (required for aggregate datasets).
*sort_by: Variables to sort the results by.
*exclude: Variables to exclude from the result.

Download the entire table with ID 142:

```{r rdbhapi, warning=FALSE, message=FALSE, results='asis'}
institusjoner <- dbh_data(142)
```


Download filtered and grouped data for table with ID 142:

```{r erasmus_plus, warning=FALSE, message=FALSE, results='asis', echo=TRUE, eval=TRUE}
library(knitr)
erasmus_plus <- dbh_data(
  142,
  filters = list(
    "\u00c5rstall" = c("top", "5"),
    Utvekslingsavtale = "ERASMUS+",
    Type = "NORSK",
    "Niv\u00e5kode" = "*"
  ),
  exclude = c("Niv\u00e5kode" = "FU"),
  group_by = "\u00c5rstall"
)
kable(head(erasmus_plus))

```



## Token facility 
To authenticate with the DBH-API, you need to provide a valid token using your` SSO ID` and `SSO Secret`. These credentials can be securely defined as environment variables in one of the following ways:
Option 1: Using `.Renviron` File
* Locate/Create the `.Renviron` File:
The `.Renviron` file is a hidden file used by R to store environment variables.
You can create or edit this file in your home directory or in your R project folder.
Open the `.Renviron` file in a text editor or within R using:

file.edit("~/.Renviron")

Add  `dbhapi_sso_id` = "your_sso_id" and `dbhapi_sso_secret` = "your_sso_secret"

 in `.Renviron` file and save the file. 

After saving, restart your R session to load the updated `.Renviron` file.

Once defined, these variables can be accessed in R using `Sys.setenv`:

```{r Renvirion access}
Sys.getenv("dbhapi_sso_id")
Sys.getenv("dbhapi_sso_secret")
```
Option 2: Using `Sys.setenv`
Alternatively, you can define the token dynamically during your R session using the `Sys.setenv` function:

```{r Sys.setenv }
Sys.setenv(dbhapi_sso_id = "your_sso_id")
Sys.setenv(dbhapi_sso_secret = "your_sso_secret")

```
## Session info
This vignette was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```
 
